### 2019-10-22 manuscript/manuscript.rmd
update manuscript

Casey W. Dunn^1^*, Benjamin Evans^2^, and Yuanning Li^1^

@@

@@

### Data wrangling

@@

### Phylogenetic analyses

@@

We thank the Yale Center for Research Computing for use of the research computing infrastructure, specificaly the Farnam cluser.

@@

Dunn *et al.* [@Dunn:2008ky] added Expressed Sequence Tag (EST) data for 29 animals. It was the first phylogenomic analysis that included ctenophores, and therefore that could test the relationships of both Ctenophora and Porifera to the rest of animals. It was the first phylogenetic analysis to recover Ctenophora as the sister group to all other animals.

@@

Philippe *et al.* 2009 [Philippe:2009hh]...

>>>>

Yuanning Li^1^, Xingxing Shen, Benjamin Evans^2^, Antonis Rokas, and Casey W. Dunn^1^*

@@

- C10 to C60 [@si2008empirical]. 10 to 60-profile mixture models as variants of the CAT model under the maximum-likelihood framework.

- Six-state amino acid recoding. Amino acids are recoded into six groups based on function to account for both compositional heterogeneity and substitution saturation. Several recoding strategies have been proposed, including Dayhoff 6-state recoding, S&R 6-state recoding, KGB 6-state recoding.

@@

### Data selection and wrangling

@@

### Phylogenetic analyses
#### Full data matrices
To investigate the phylogenetic hypotheses and distribution of phylogenetic signal in studies aiming to elucidate the root of animal position, we considered 16 data matrices from recent studies that were constructed from transcriptomic or genomic data (Table). Because different choices of substitution models could largely influence phylogenetic inference of the placement of the root of animal (e.g. site heterogeneous CAT model and site homogeneous model), we first investigated four different sets of substitution models in IQtree [@nguyen2014iq], including C10 to C60 profile mixture models as variants of the CAT model in ML framework. Phylogenomic analyses of all the datasets were first conducted under WAG+G, GTR+G, associated best-fit substitution model (C10-C60 model were included for model comparison via -madd option) with ModelFinder [@kalyaanamoorthy2017modelfinder] and Poisson + C60 models using IQtree v [@nguyen2014iq]. Nodal support was assessed with 1000 ultrafast bootstrap replicates for each analysis.

#### Outgroup taxa sampling with C10-C60 and CAT models
Because different choices of outgroups could also affect phylogenetic inference as suggested in previous analyses, we parsed the full data matrices into three different types of outgroups: Choanimalia , Holozoa and Opisthokonta. These datasets include the same set of genes but differ in the composition of outgroup species. Choanozoa only includes choanofagellates as outgroup; Holozoa also includes more distantly related holozoans; Opistokonta also includes Fungi. For each Choanozoa data matrice, both C10-C60 models in ML and CAT models in PhyloBayes were conducted. The maximum likelihood analysis was performed using the best-fit substitution model identified as above and nodal support was assessed with 1000 ultrafast bootstrap replicates using IQtree. Moreover, bayesian inference with the site-heterogeneous CAT-Poisson substitution model was done with PhyloBayes MPI. To minimize computational burden, CAT-GTR models were not in this study.

For results of choanozoa matrices indicated a strong support that sponges are the sister group to the remaining Metazoa using CAT model, bayesian inference with CAT-Poisson model was also conducted to Holozoa and Opisthokonta data matrices with the same settings as above.  For CAT-poisson analyses on all the data matrices, two independent chains were sampled every generation. Tracer plots of MCMC runs were visually inspected in Tracer v1.6 to assess stationarity and appropriate burn-in. Chains were considered to have reached convergence when the maxdiff statistic among chains was below 0.3 (as measured by bpcomp) and effective sample size > 50 for each parameter (as measured by tracecomp). A 50% majority‐rule consensus tree was computed with bpcomp, and nodal support was estimated by posterior probability. All analyses in PhyloBayes were run for at least one month computational time.

#### phylogenetic signal in conflicted dataset
To investigate the distribution of phylogenetic signal in data matrices, we considered four data matrices from four studies that had different topology between ML and BI using CAT model in our reanalysis, including Philippe2008, Ryan2013, and Whelan_2017 data matrices.We examined two hypotheses: Ctenophora-sister and Porifera-sister to rest of metazoans, under both ML and BI frameworks with different outgroup schemes (Choanozoa and Opisthokonta). For ML analysis in each dataset, site-wise likelihood scores ere inferred for both hypotheses using IQtree (option -g) with the same best-fit model identified above. The two different
phylogenetic hypotheses passed to IQtree (via -z) were the corresponding tree that the ctenophore as the sister lineage tree and the corresponding
tree that was modified to have sponges as the sister to all other metazoans. The constraint trees were conducted by R package.  The numbers of genes and sites supporting each hypothesis were calculated with RAxML output and Perl scripts from Shen et al. For BI analysis, we only considered the Whelan_2017 dataset.

@@

We thank the Yale Center for Research Computing for use of the research computing infrastructure, specificaly the Farnam cluster.

@@

Dunn *et al.* [@Dunn:2008ky] added Expressed Sequence Tag (EST) data for 29 animals. It was the first phylogenomic analysis that included ctenophores, and therefore that could test the relationships of both Ctenophora and Porifera to the rest of animals. It was also the first phylogenetic analysis to recover Ctenophora as the sister group to all other animals.

@@

Philippe *et al.* 2009 [Philippe:2009hh] assembled a 128 EST dataset for 55 species to explore phylogenetic relationship of early diverging animals.
